
---
title: "SV table with HOMVAR HOMEREF ect. AF=.5 or not"
output:
  word_document: default
  html_document: default
  pdf_document: default
date: "2023-11-21"
---
```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE)

setwd("/Users/mulch/Documents/R_working_directory")

# Set the root directory for notebook chunks
knitr::opts_knit$set(root.dir = "/Users/mulch/Documents/R_working_directory")

library(dplyr)
library(ggplot2)


```

load the table with excess het homvar and hom ref
see how many of each, and stratily by technology and by if AF=.5 or not
```{r}
SV_table_w_excesshet <- read.table('svData.table.exhet.txt', sep = '\t', header = TRUE)

#  first row
print(SV_table_w_excesshet[1,])

# column numbers and their names
for (i in 1:ncol(SV_table_w_excesshet)) {
  cat("Column", i, ": ", colnames(SV_table_w_excesshet)[i], "\n")
}

View(SV_table_w_excesshet)
```

add a col for AF=.5 or not this is TRUE or FALSE 
```{r}
SV_table_w_excesshet$AF_0.5 <- SV_table_w_excesshet$AF == 0.5

#also add a col for technology for if ID col contains CLR or not 

#parsing the ID colSV_table_w_excesshet
SV_table_w_excesshet <- SV_table_w_excesshet %>%
  mutate(technology = case_when(
    grepl("clr", ID) & grepl("ccs", ID) ~ "BOTH",
    grepl("clr", ID) ~ "CLR",
    grepl("ccs", ID) ~ "CCS",
    TRUE ~ "OTHER"
  ))

```

Here is some info about some of the categories of interest
HET (count of het genotypes)
HOM-REF (count of homozygous reference genotypes)
HOM-VAR (count of homozygous variant genotypes) 


Lets look as homvar and homref by svtype as a first look
 
```{r}

#percent for a few categories
homvar_homref_svtype <- SV_table_w_excesshet  %>%
  group_by(SVTYPE, HOM.REF, HOM.VAR) %>%
  summarise(Percent = n() / nrow(SV_table_w_excesshet) * 100)


#  SVTYPE homvar
plot_counts_homvar_SVTYPE <- ggplot(homvar_homref_svtype, aes(x = HOM.VAR, y = Percent, fill = SVTYPE)) +
  geom_bar(stat = "identity", position = position_dodge(width = 0.9)) +
  labs(
    x = "homvar",
    y = "Percent of total SVs",
    fill = "SVTYPE",
    title = "percent total variants by homvar",
    subtitle = "Grouped Bar Plot"
  ) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

#  SVTYPE and homref
plot_counts_homref_SVTYPE <- ggplot(homvar_homref_svtype, aes(x = HOM.REF, y = Percent, fill = SVTYPE)) +
  geom_bar(stat = "identity", position = position_dodge(width = 0.9)) +
  labs(
    x = "homref",
    y = "Percent of total SVs",
    fill = "SVTYPE",
    title = "percent total variants by homref",
    subtitle = "Grouped Bar Plot"
  ) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

print(plot_counts_homvar_SVTYPE)
print(plot_counts_homref_SVTYPE)

# Fraction for each category
homvar_homref_svtype_fraction <- SV_table_w_excesshet %>%
  group_by(SVTYPE, HOM.REF, HOM.VAR, HET) %>%
  summarise(Fraction = n() / nrow(SV_table_w_excesshet))

 # SVTYPE and het
plot_fraction_het_SVTYPE <- ggplot(homvar_homref_svtype_fraction, aes(x = HET, y = Fraction, fill = SVTYPE)) +
  geom_bar(stat = "identity", position = "fill") +
  labs(
    x = "HET",
    y = "Fraction of total SVs",
    fill = "SVTYPE",
    title = "Fraction total variants by het",
    subtitle = "Grouped Bar Plot"
  ) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

# SVTYPE and homvar
plot_fraction_homvar_SVTYPE <- ggplot(homvar_homref_svtype_fraction, aes(x = HOM.VAR, y = Fraction, fill = SVTYPE)) +
  geom_bar(stat = "identity", position = "fill") +
  labs(
    x = "homvar",
    y = "Fraction of total SVs",
    fill = "SVTYPE",
    title = "Fraction total variants by homvar",
    subtitle = "Grouped Bar Plot"
  ) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

# SVTYPE and homref
plot_fraction_homref_SVTYPE <- ggplot(homvar_homref_svtype_fraction, aes(x = HOM.REF, y = Fraction, fill = SVTYPE)) +
  geom_bar(stat = "identity", position = "fill") +
  labs(
    x = "homref",
    y = "Fraction of total SVs",
    fill = "SVTYPE",
    title = "Fraction total variants by homref",
    subtitle = "Grouped Bar Plot"
  ) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

print(plot_fraction_homvar_SVTYPE)
print(plot_fraction_homref_SVTYPE)
print(plot_fraction_het_SVTYPE)

```
```{r}
# Fraction for each category
het_svtype_fraction <- SV_table_w_excesshet %>%
  group_by(technology, HET, AF_0.5, NCALLED) %>%
  summarise(Fraction = n() / nrow(SV_table_w_excesshet))

# SVTYPE and homref
plot_fraction_het <- ggplot(het_svtype_fraction, aes(x = HET, y = Fraction, fill = AF_0.5)) +
  geom_bar(stat = "identity", position = "fill") +
  labs(
    x = "HET",
    y = "Fraction of total SVs",
    fill = "AF=.5",
    title = "Fraction total variants by HET",
    subtitle = "Grouped Bar Plot"
  ) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

print(plot_fraction_het)

plot_fraction_tec <- ggplot(het_svtype_fraction, aes(x = HET, y = Fraction, fill = technology)) +
  geom_bar(stat = "identity", position = "fill") +
  labs(
    x = "HET",
    y = "Fraction of total SVs",
    fill = "technology",
    title = "Fraction total variants by HET",
    subtitle = "Grouped Bar Plot"
  ) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

print(plot_fraction_tec)
#subset of AF=.5
het_svtype_fraction_subset <- SV_table_w_excesshet %>%
  filter(AF_0.5 == TRUE) %>%
  group_by(technology, HET, NCALLED) %>%
  summarise(Fraction_AF.5 = n() / nrow(het_svtype_fraction_subset))


plot_counts_het_SVTYPE_subset <- ggplot(het_svtype_fraction_subset, aes(x = HET, y = Fraction_AF.5, fill = technology)) +
  geom_bar(stat = "identity", position = "fill") +
  labs(
    x = "het",
    y = "Fraction of total SVs where AF=.5",
    fill = "technology",
    title = "Fraction total variants by het (AF_0.5 subset)",
    subtitle = "Grouped Bar Plot"
  ) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

print(plot_counts_het_SVTYPE_subset)
```
```{r}
#fraction of het
het_AF.5_fraction <- SV_table_w_excesshet %>%
  group_by(technology, HET, NCALLED, AF_0.5) %>%
  summarise(Fraction_Het_Ncalled = (HET) / (NCALLED))

plot_het <- ggplot(het_AF.5_fraction, aes(x = technology, y = Fraction_Het_Ncalled)) +
  geom_bar(stat = "identity", position = "fill") +
  labs(
    x = "technology",
    y = "Fraction of HET / NCALLED",
    title = "Fraction HET / NCALLED ",
    subtitle = "Grouped Bar Plot"
  ) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

print(plot_het)

#do a geom_density or geom_violin where x axis is fraction HET make y axis the technology and
plot_het_violin <- ggplot(het_AF.5_fraction, aes(x = technology, y = Fraction_Het_Ncalled)) +
  geom_violin(scale = "width", position = "fill", trim = FALSE) +
  labs(
    x = "Technology",
    y = "Fraction of HET / NCALLED",
    title = "Fraction HET / NCALLED",
    subtitle = "Violin Plot"
  ) +
  theme_minimal() +
  facet_wrap(~ technology, scales = "free")

print(plot_het_violin)

plot_het_density_facet <- ggplot(het_AF.5_fraction, aes(x = Fraction_Het_Ncalled)) +
  geom_density(alpha = 0.5) +
  labs(
    x = "Fraction of HET / NCALLED",
    title = "Density Plot of Fraction HET / NCALLED",
    subtitle = "Density Plot"
  ) +
  scale_fill_discrete(name = "AF_0.5") +
  theme_minimal() +
  facet_wrap(~ technology, scales = "free")

print(plot_het_density_facet)
```

```{r}
#fraction of het
het_AF.5_fraction <- SV_table_w_excesshet %>%
  group_by(technology, HET, NCALLED, AF_0.5) %>%
  summarise(Fraction_Het_Ncalled = (HET) / (NCALLED))

plot_het_AF.5 <- ggplot(het_AF.5_fraction, aes(x = technology, y = Fraction_Het_Ncalled, fill = AF_0.5)) +
  geom_bar(stat = "identity", position = "fill") +
  labs(
    x = "technology",
    y = "Fraction of HET / NCALLED",
    fill = "AF_0.5",
    title = "Fraction HET / NCALLED ",
    subtitle = "Grouped Bar Plot"
  ) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

print(plot_het_AF.5)

#do a geom_density or geom_violin where x axis is fraction HET make y axis the technology and facet based on AF_0.5
plot_het_AF.5_violin <- ggplot(het_AF.5_fraction, aes(x = technology, y = Fraction_Het_Ncalled, fill = as.factor(AF_0.5))) +
  geom_violin(scale = "width", position = "fill", trim = FALSE) +
  labs(
    x = "Technology",
    y = "Fraction of HET / NCALLED",
    fill = "AF_0.5",
    title = "Fraction HET / NCALLED",
    subtitle = "Violin Plot"
  ) +
  scale_fill_discrete(name = "AF_0.5") +
  theme_minimal() +
  facet_wrap(~ technology, scales = "free")

print(plot_het_AF.5_violin)

plot_het_AF.5_density_facet <- ggplot(het_AF.5_fraction, aes(x = Fraction_Het_Ncalled, fill = as.factor(AF_0.5))) +
  geom_density(alpha = 0.5) +
  labs(
    x = "Fraction of HET / NCALLED",
    fill = "AF_0.5",
    title = "Density Plot of Fraction HET / NCALLED",
    subtitle = "Density Plot"
  ) +
  scale_fill_discrete(name = "AF_0.5") +
  theme_minimal() +
  facet_wrap(~ technology, scales = "free")

print(plot_het_AF.5_density_facet)


```

```{r}
#ok now strat by if AF=.5 or not and technology

# percentage of total SVs for a few variables
tech_af_0.5_count <- SV_table_w_excesshet  %>%
  group_by(technology, AF_0.5, HOM.VAR, HOM.REF) %>%
  summarise(Count = n())

tech_af_0.5 <- SV_table_w_excesshet %>%
  group_by(technology, AF_0.5, HOM.VAR, HOM.REF) %>%
  summarise(Percent = n() / nrow(SV_table_w_excesshet) * 100)

#  SVTYPE and af.5 for
plot_percent_AF.5 <- ggplot(tech_af_0.5, aes(x = technology, y = Percent, fill = AF_0.5)) +
  geom_bar(stat = "identity", position = position_dodge(width = 0.9)) +
  labs(
    x = "technology",
    y = "Percent of Total SVs",
    fill = "AF is .5",
    title = "percent of total SVs that are AF=.05",
    subtitle = "AF is .5 by technology"
  ) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

#  SVTYPE and af.5 
plot_percent_AF.5_count <- ggplot(tech_af_0.5_count, aes(x = technology, y = Count, fill = AF_0.5)) +
  geom_bar(stat = "identity", position = position_dodge(width = 0.9)) +
  labs(
    x = "technology",
    y = "Count",
    fill = "AF is .5",
    title = "count of total SVs that are AF=.05",
    subtitle = "AF is .5 by technology"
  ) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

print(plot_percent_AF.5_count)
print(plot_percent_AF.5)

#fraction vs percent of total
tech_af_0.5_fraction <- SV_table_w_excesshet %>%
  group_by(technology, AF_0.5, HOM.VAR, HOM.REF) %>%
  summarise(Fraction = n() / nrow(SV_table_w_excesshet))

### add in stacked bar for AF=.5 or not( out of the total)
plot_fraction_AF_0.5_total <- ggplot(tech_af_0.5_fraction, aes(x = technology, y = Fraction, fill = AF_0.5)) +
  geom_bar(stat = "identity", position = "stack") +
  labs(
    x = "Technology",
    y = "Fraction of Total SVs",
    fill = "AF is 0.5",
    title = "Fraction of Total SVs with AF=0.5",
    subtitle = "AF is 0.5 by Technology"
  ) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

print(plot_fraction_AF_0.5_total)


```



below is heatmap example plot in case this can be useful
```{r}

#Heatmaps?? hom var and hom ref and how thay may be coreleted

#AF
heatmap_homvar_homref <- ggplot(SV_table_w_excesshet, aes(x = HOM.VAR, y = HOM.REF, fill = AF)) +
  geom_tile() +
  scale_fill_gradient(low = "white", high = "blue") +  # Adjust color scale
  labs(
    title = "Heatmap of AF by homvar and homref",
    x = "homvar",
    y = "homref"
  ) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),  
    legend.position = "right"  
  )

#svlenth
 heatmap_homvar_homref_svlen <- ggplot(SV_table_w_excesshet, aes(x = HOM.VAR, y = HOM.REF, fill = SVLEN)) +
  geom_tile() +
  scale_fill_gradient(low = "white", high = "blue") +  # Adjust color scale
  labs(
    title = "Heatmap of Sv Length by homvar and homref",
    x = "homvar",
    y = "homref"
  ) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),  
    legend.position = "right"  
  )

print(heatmap_homvar_homref_svlen)
print(heatmap_homvar_homref)

```

```{r}
#plotting homvar and hom ref a few ways
ggplot(tech_af_0.5, aes(x = technology, y = Percent, fill = factor(HOM.VAR))) +
  geom_bar(stat = "identity", position = "stack") +
  labs(
    x = "Technology",
    y = "Percentage",
    fill = "HOM.VAR",
    title = "Percentage of HOM.VAR by Technology"
  ) +
  theme_minimal()

ggplot(SV_table_w_excesshet, aes(x = technology, y = NCALLED, fill = factor(HOM.VAR))) +
  geom_bar(stat = "identity", position = "fill") +
  labs(
    x = "Technology",
    y = "N called",
    fill = "HOM.VAR",
    title = "Stacked NCALLED of HOM.VAR by Technology"
  ) +
  theme_minimal()
#next two should get at Sams points:
ggplot(SV_table_w_excesshet, aes(x = technology, y = NCALLED, fill = factor(HET))) +
  geom_bar(stat = "identity", position = "fill") +
  labs(
    x = "Technology",
    y = "N called",
    fill = "HET",
    title = "Stacked NCALLED of HET by Technology"
  ) +
  theme_minimal()
ggplot(SV_table_w_excesshet, aes(x = HET, y = NCALLED, fill = factor(AF_0.5))) +
  geom_bar(stat = "identity", position = "fill") +
  labs(
    x = "HET",
    y = "N called",
    fill = "AF_0.5",
    title = "Stacked HET AF_0.5 for N called"
  ) +
  theme_minimal()

ggplot(tech_af_0.5, aes(x = technology, y = Percent, fill = factor(AF_0.5), color = factor(HOM.REF))) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(
    x = "Technology",
    y = "Percentage",
    fill = "AF=0.5",
    color = "HOM.REF",
    title = "Percentage of AF=0.5 by Technology"
  ) +
  theme_minimal()


```
Summerize by ncalled
```{r}
# percentage of total SVs for a few variables
tech_ncalled_count <- SV_table_w_excesshet  %>%
  group_by(technology, AF_0.5, HOM.VAR, HOM.REF, NCALLED) %>%
  summarise(Count = n())

tech_ncalled <- SV_table_w_excesshet %>%
  group_by(technology, AF_0.5, HOM.VAR, HOM.REF, NCALLED) %>%
  summarise(Percent = n() / nrow(SV_table_w_excesshet) * 100)

ggplot(tech_ncalled, aes(x = technology, y = Percent, fill = factor(NCALLED), color = factor(HOM.VAR))) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(
    x = "Technology",
    y = "Percentage",
    fill = "N called",
    color = "technology",
    title = "Percentage of Ncalled by Technology"
  ) +
  theme_minimal()

ggplot(tech_ncalled_count, aes(x = HOM.VAR, y = NCALLED, fill = factor(HOM.REF), color = factor(technology))) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(
    x = "HOMVAR",
    y = "NCALLED",
    fill = "N called ",
    color = "tech",
    title = "HOMVAR by Ncalled and Technology"
  ) +
  theme_minimal()

ggplot(tech_ncalled_count, aes(x = HOM.REF, y = NCALLED , color = factor(technology))) +
  geom_bar(stat = "identity", position = "dodge", alpha = 0.8) +
  labs(
    x = "HOMREF",
    y = "NCALLED",
    fill = "N called ",
    color = "technology",
    title = "Ncalled by Technology for HOMREF"
  ) +
  theme_minimal()

```


here are the bins for AF and Length that are used in other plots- incorporate these?
```{r}
# Define brackets 
bins_length <- c(-1000000, -100000, -10000, -7000, -5000, -1000, -350, -250, -100, -10, 10, 100, 250, 350, 1000, 5000, 7000, 10000, 100000, 1000000)
tags <- c("-100kb - -1000kb", "-10kb - -100kb", "-7kb - -10kb", "-5kb - -7kb", "-1kb - -5kb", "-350bp - -1kb", "-250bp - -350bp", "-100bp - -250bp", "-10bp - -100bp", "<10bp", "10bp - 100bp", "100bp - 250bp", "250bp - 350bp", "350bp - 1kb", "1kb - 5kb", "5kb - 7kb", "7kb - 10kb", "10kb - 100kb", "100kb - 1000kb")

# Bin the SVLEN values
bins_svlen <- df %>%
  mutate(SVLEN_Bracket = cut(SVLEN, breaks = breaks, labels = tags, include.lowest = TRUE))

```


```{r}
#AF bins
bins_AF <- cut(df$AF, 
            breaks = c(0, 0.15, 0.33, 0.9, 1), 
            labels = c("Ultra Rare", "Rare", "Common", "AF=1"),
            include.lowest = TRUE)



```

